### 2020-12-03 analysis_and_paper/manuscript.Rmd
Corrects some paths and knitts analysis for experiment 1 and supplements.
Adds updated files following IPA at PS.

author: 
  - name          : Tobias Heycke
    affiliation   : 1, 2
    corresponding : yes    
    address       : "P.O. 122155, 68072 Mannheim, Germany"
    email         : "tobias.heycke@gesis.org"
  - name          : Frederik Aust
    affiliation   : 1
  - name          : Mahzarin R. Banaji
    affiliation   : 3
  - name          : Pieter Van Dessel
    affiliation   : 5
  - name          : Xiaoqing Hu
    affiliation   : 6    
  - name          : Congjiao Jiang
    affiliation   : 4    
  - name          : Benedek Kurdi
    affiliation   : 3    
  - name          : Robert Rydell
    affiliation   : 7
  - name          : Lisa Spitzer
    affiliation   : 1
  - name          : Christoph Stahl
    affiliation   : 1
  - name          : Christine Vitiello
    affiliation   : 4
  - name          : Jan De Houwer
    affiliation   : 5

affiliation:
  - id            : 1
    institution   : University of Cologne
  - id            : 2
    institution   : GESIS - Leibniz Institute for the Social Sciences
  - id            : 3
    institution   : Harvard University 
  - id            : 4
    institution   : University of Florida 
  - id            : 5
    institution   : Ghent University 
  - id            : 6
    institution   : The University of Hong Kong 
  - id            : 7
    institution   : Indiana University

abstract: |
  Findings of dissociations between implicit (i.e., automatic) and explicit (i.e., non-automatic) evaluations that are based on distinct associative (i.e., co-occurrence based) and propositional (i.e., rule-based) learning procedures have fueled the dominance of dual-process theories of evaluative learning for decades.  Arguably the most influential evidence has been found in a study by Rydell, McConnell, Mackie, and Strain (2006) in which participants learned about a person named Bob. It was observed that implicit evaluations reflected the valence of brief pairings of valenced words with the image of Bob whereas explicit evaluations reflected the (opposite) valence of the behavioral statements that were instructed to be characteristic of Bob. A recent study by Heycke and colleagues (2018) was unable to reproduce this data pattern independently. Given the theoretical importance of the findings by Rydell and colleagues, we present a series of additional replication attempts conducted by an international collective of researchers including the first author of the original finding.

authornote: |
  All data, analysis scripts and materials can be found at https://osf.io/8m3xb/

keywords          : "evaluative conditioning, subliminal influence, implicit learning, replication"
#wordcount         : "X"

@@

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no
draft             : no

@@

library("afex")
library("emmeans")
library("BayesFactor")
library("MBESS")

@@

source("modal_frame_rate.R")
```

@@

<!--
Guideline:
Authors of a PDR must make a convincing case that the replication will make a valuable contribution to understanding a phenomenon and/or theory of broad current interest to psychologists. As with all Psychological Science submissions, the primary criterion is general theoretical significance. Direct replications should reproduce the original methods and procedures as closely as possible, with the goal of measuring the same effect under essentially (but not necessarily superficially) the same conditions as in the original study

Word limit: 
Intro + Discussion + notes + acknowledgements + appendices = 1500

Max Figures/tables:
3

Max references:
30
-->

Evaluative conditioning (EC) refers to the robust finding of a change in liking of a stimulus, referred to as conditioned stimulus (CS), that is due to its pairing with a stimulus of positive or negative valence, referred to as unconditioned stimulus (US) [@dehouwerConceptualTheoreticalAnalysis2007; @hofmannEvaluativeConditioningHumans2010]. 
Dual-process theories postulate that EC can depend both on a mechanism via which co-occurrences between stimuli result in the automatic formation of mental associations as well as a second mechanism via which propositional beliefs are formed [e.g., @gawronskiAssociativePropositionalProcesses2006; @rydellUnderstandingImplicitExplicit2006].
Though dual-process theories have had a huge impact, it has been argued that there is actually little evidence for automatic association formation as a distinct mechanism for EC [see @corneilleAssociativeAttitudeLearning2018a].
This conclusion is in line with propositional single-process views [e.g., @mitchellPropositionalNatureHuman2009] which postulate that all associative (evaluative) learning (including EC) is based on propositional processes.

One finding, however, is particularly difficult to reconcile with propositional single-process accounts: 
@rydell_two_2006 reported that participants exhibited conflicting explicit and implicit evaluations (e.g., positive explicit and negative implicit evaluations) of a single target person after the presentation of rapid stimulus-stimulus pairings and behavioral statements that implied the opposite valence. 
In this study, participants engaged in a learning task about a person named Bob in which they determined whether Bob was a good or bad person by guessing whether positive and negative behavioral statements were characteristic of Bob. 
They received feedback about the accuracy of each of their guesses.
Unbeknownst to the participants, on each trial, a valent word (US) was presented briefly (25 ms) prior to the presentation of a picture of Bob and the behavioral statement. 
Critically, the valence of the US words was always opposite to the behavioral information (e.g., USs were always positive when negative behavioral statements were characteristic of Bob).
Following the learning task, explicit evaluations of Bob, assessed with a trait rating task, reflected the valence of the learned behavioral information, while implicit evaluations, assessed with an Implicit Association Test [IAT: @greenwaldMeasuringIndividualDifferences1998] reflected the valence of the USs.
These results were conceptually replicated by the original authors using similar methods [@rydellConsequencesDiscrepantExplicit2008; @rydellUnderstandingImplicitExplicit2006].
The @rydell_two_2006 article has been referred to as providing strong evidence for numerous dual-process theories [@gawronskiAssociativePropositionalEvaluation2011; @rydellUnderstandingImplicitExplicit2006] and has been cited more than 320 times overall (Google Scholar, 10.15.2019).

Judging from its impact, the reported effect is generally assumed to be robust and replicable.
A recent study, composed of two experiments, however, could not replicate the dissociative effect [@heycke_two_2018].
Instead, both implicit and explicit evaluations consistently reflected the valence of the learned behavioral information.
At present, it remains unclear whether these unsuccessful attempts call into question the replicability of the original study or point towards boundary conditions.
This ambiguity is due to the modifications made to the original materials and procedure in the replication studies: First, some words were replaced to ensure that no words presented in the learning phase served as distractors in the US memory assessment at the end of the study [see @heycke_two_2018].
Second, one experiment used German translations of the verbal material (rather than the original English material).
Third, the other experiment, while using the original material, was conducted with a modified procedure: Because  a subset of participants correctly identified USs after the learning task in the first experiment, the presentation duration of USs was reduced from 25 ms to 16 ms to ensure at-chance post-learning US recognition.

In the current study we will rigorously test the replicability of the original findings reported by @rydell_two_2006 by closely adhering to the original procedure.
To ensure its informativeness, the current replication attempt was a joint effort that involved researchers from different labs that all have expertise in studying evaluative conditioning and implicit measures.
Moreover, both the first author of the original study and authors of the unsuccessful replication attempts collaborated on the studies presented here.
To explore the generality of our results, data were collected in multiple countries and languages.
In a first, already concluded,  experiment we have conducted a close replication in three different locations. 
In a second experiment, for which the data has not been collected yet, we will use the knowledge gained from the first experiment to adjust the procedure to closely replicate the psychological conditions of the original study.

@@

  # Recode, calculate indices, and filter data
  otm1_eval <- otm1_eval %>%
    mutate_at(vars(starts_with("Eval")), scale) %>%
    select(starts_with("Eval")) %>%
    rowwise %>%
    do(data.frame(Eval = sum(unlist(.)))) %>% 
    bind_cols(otm1_eval, .) %>%
    mutate(
      ParticipantNumber = factor(ParticipantNumber)
      , ValenceBlock = ifelse(ValenceBlock == 1, "Positive-negative", "Negative-positive") %>% factor
      , MeasureOrder = ifelse(MeasureOrder == 1, "Implicit-explicit", "Explicit-implicit") %>% factor
      , Block = factor(Block)
    )

@@

As the results reported by @heycke_two_2018 could have been the result of the modifications to the experimental procedure, we conducted a replication study using the unmodified experimental procedure of the original study.

@@

Materials and procedure were inspected by the first author of the original study to ensure that they closely corresponded to the original.
The experiment was preregistered (https://osf.io/xe8au/) and data were collected in Germany, Belgium, and the United States.
All data files, materials and analysis files are available at https://osf.io/8m3xb/.

@@

The experimental procedure consisted of three components: learning task, evaluation task, and US recognition task.

As in the original study, the learning task was a modified version of the evaluative learning paradigm by @kerpelmanPartialReinforcementEffects1971. 
We briefly presented a US followed by a longer presentation of a picture of Bob together with a statement that described the behavior of Bob.
Presentation durations differed across labs due to the availability of different refresh rates of the CRT monitors (85 Hz in the United States and 75 Hz in Belgium and Germany).
In the following we will describe the setup of a trial with the presentation durations at a 75 Hz-refresh rate; deviating durations for a 85 Hz-refresh rate are given in brackets.

On each trial, a central fixation cross was displayed for 200 ms followed by a US displayed for 27 ms (24 ms; 2 frames).
USs were immediately replaced by the picture of Bob, which served as a backward mask.
The picture of Bob was presented in the center of the screen for 253 ms (247 ms) before the behavioral information was added underneath.
The screen background was black and text was colored white and set in Times New Roman font. 
Participants’ task was to press the "c" (= "characteristic") or "u" (= "uncharacteristic") key to judge whether the behavioral information was characteristic or uncharacteristic of Bob, which was displayed on the screen during the judgment phase. 
After every judgment, the image of Bob, the behavioral statement, and the key labels were replaced with either the word "Correct" displayed in green letters or the word "False" in red letters, displayed for 5000 ms. 
Each trial ended with a blank screen presented for 1000 ms.

As US valence was manipulated within participants, they completed two 100-trial-blocks of the learning task.
Each block consisted of trials with either only positive or negative USs and the order of USs was randomized. 
In blocks with positive USs, positive behavioral information was always uncharacteristic of Bob and negative information was characteristic. 
These contingencies were reversed in the blocks with negative USs. 
We used 10 positive and 10 negative words as USs each of which was presented 10 times. 
For behavioral information, we used 100 positive and 100 negative statements. 
50 positive and 50 negative statements were randomly selected for the first block, the remaining statements were assigned to the second block. 
The order of USs and behavioral information was randomized for each participant anew, whereas the order of blocks was counterbalanced across participants. 
A different picture of Bob was randomly selected from six pictures of white males for each participant. 
The remaining five images were used in the implicit evaluation measure (see below). 
All materials were taken from the original study, with the sole exception that US words, behavioral information, and instructions were translated to German and Dutch for use in Germany and Belgium.

After each block, participants completed implicit and explicit evaluation measures. 
As in the original study, the order of the measures was the same for both blocks but counterbalanced across participants.

The explicit evaluation consisted of three parts, each presented on a single screen:
First, participants rated Bob's likableness on a 9-point slider with the anchors labelled *Very Unlikable* and *Very Likable*.
Next, again using 9-point sliders, they judged Bob on the dimensions *Bad--Good*, *Mean--Pleasant*, *Disagreeable--Agreeable*, *Uncaring--Caring*, and *Cruel--Kind*. 
Finally, they judged Bob on a 'feeling thermometer' by entering a number between 0 (*Extremely unfavorable*) and 100 (*Extremely favorable*).
Deviating from the original protocol, we collected explicit evaluations as part of the computer task rather than using a paper-pencil questionnaire.

We assessed the implicit evaluations using the IAT described by @rydell_two_2006. 
Participants initially completed two types of training blocks with 20 trials each to familiarize themselves with the task. 
In one block, images of Bob and other white men had to be classified as Bob vs. not-Bob; in another block, positive and negative words had to be classified as positive vs. negative. 
In a subsequent critical block with 40 trials we intermixed the two classification tasks: 
Participants used one key to respond to both the images of Bob and negative words; they used another key to respond to images of other white men and positive words. 
After the first critical block, participants completed another training block with 20 trials of Bob vs. not-Bob with reversed key position and afterwards a second critical block with 40 trials with the reversed key mapping compared to the first critical block. 
It was counterbalanced whether participants completed the IAT with the key mappings described above or with key mappings in reversed order [for a detailed description see @heycke_two_2018, p. 1712].

After completing the explicit and implicit evaluation measures, participants completed the second learning block and again completed implicit and explicit evaluations.
Following the second set of evaluations, participants completed a surprise US recognition task.
We presented 40 words in random order on a computer screen.
Half of the words were the briefly presented USs from the learning task, the other half were new distractor words.
We informed participants that 20 words were presented briefly during the learning task, asked them to select the briefly presented words from the list, and encouraged them to guess if they did not know the correct answer.
Participants could only proceed with the experiment once they had selected exactly 20 words.

The experiment ended with a demographic questionnaire (age, field of study/profession, gender, goal of the experiment, and comments).
Our procedure was identical to the original procedure, with the exception that participants completed explicit evaluation and US recognition tasks at the computer rather than using paper and pencil.
In Belgium and Germany, we furthermore used Dutch and German translations of the original material.
The procedure took approximately 50 minutes to complete.

@@

We set out to collect $N = 50$ participants at each location [@rydell_two_2006].
For each evaluation measure, the data from all three labs ($N = 150$) provides 95% power to observe two-way interaction effects as small as $f = 0.15$ ($d = 0.30$) and contrasts between blocks of the learning task as small as $d_z = 0.27$ [$\alpha = \beta = .05$, repeated-measures correlation $r = .5$, @nosek_implicit_2007].
Our design was, thus, adequately powered to detect effects half the size of those reported by @rydell_two_2006.
We recruited `r otm1_n + otm1_n_excluded` participants (aged `r paste(range(otm1_mem$Age), collapse = "-")` years, $M = `r mean(otm1_mem$Age)`$; `r sum(grepl("female", otm1_mem$Sex)) / nrow(otm1_mem) * 100`% female, `r printnum(sum(grepl("nonbinary", otm1_mem$Sex)) / nrow(otm1_mem) * 100)`% nonbinary; see supplementary online material (SOM) for details); `r printnum(otm1_n_excluded, numerals = otm1_n_excluded > 10, capitalize = TRUE)` participants were excluded due to technical failure<!-- or noncompliance with instructions-->.
Hence, the following results are based on data from `r otm1_n` participants.
We compensated all participants with either € 8/10 (Cologne/Ghent), or partial course credit (Cologne/Harvard).

>>>>

author: 
  - name          : Tobias Heycke
    affiliation   : "1,2"
    corresponding : yes    
    address       : "P.O. 122155, 68072 Mannheim, Germany"
    email         : "tobias.heycke@gmail.com"
    equal_contrib : yes
  - name          : Frederik Aust
    affiliation   : "1,9"
    equal_contrib : yes
  - name          : Mahzarin R. Banaji
    affiliation   : 3
  - name          : Jeremy Cone
    affiliation   : 8
  - name          : Pieter Van Dessel
    affiliation   : 5
  - name          : Melissa J. Ferguson
    affiliation   : 10
  - name          : Xiaoqing Hu
    affiliation   : 6    
  - name          : Congjiao Jiang
    affiliation   : 4    
  - name          : Benedek Kurdi
    affiliation   : "3,10"
  - name          : Robert Rydell
    affiliation   : 7
  - name          : Lisa Spitzer
    affiliation   : 1
  - name          : Christoph Stahl
    affiliation   : 1
  - name          : Christine Vitiello
    affiliation   : 4
  - name          : Jan De Houwer
    affiliation   : 5

affiliation:
  - id            : 1
    institution   : University of Cologne
  - id            : 2
    institution   : GESIS - Leibniz Institute for the Social Sciences
  - id            : 3
    institution   : Harvard University 
  - id            : 4
    institution   : University of Florida 
  - id            : 5
    institution   : Ghent University 
  - id            : 6
    institution   : The University of Hong Kong 
  - id            : 7
    institution   : Indiana University
  - id            : 8
    institution   : Williams College
  - id            : 9
    institution   : University of Amsterdam
  - id            : 10
    institution   : Yale University

abstract: "(ref:abstract)"

authornote: |
  All data, analysis scripts and materials are available at https://osf.io/8m3xb/; the supplementary online material (SOM) is available at https://osf.io/8w9bd/.

note: |
  \clearpage

keywords          : "evaluative learning, subliminal influence, implicit learning, replication"
#wordcount         : "X"

@@

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no
draft             : no
quote_labels      : yes

link-citations    : yes
linkcolor         : blue
urlcolor          : blue

@@

library("afex")
library("emmeans")
library("BayesFactor")
library("MBESS")
library("effectsize")
library("Superpower")

@@

source("modal_frame_rate.R")
source("effect_sizes.R")
source("apa_print_ps.R")
```

@@

```{r load-sample-sizes}
man_cache <- "manuscript_cache/latex/"
otm1_n <- knitr::load_cache(label = "otm1-exclusion", object = "otm1_n", path = man_cache)
otm2_n <- knitr::load_cache(label = "otm2-exclusion", object = "otm2_n", notfound = "TBD", path = man_cache)
```

(ref:abstract) Several dual-process theories of evaluative learning posit two distinct implicit (or automatic) and explicit (or controlled) evaluative learning processes. As such, one may like a person explicitly but simultaneously dislike them implicitly. Dissociations between direct measures (e.g., Likert scales), reflecting explicit evaluations, and indirect measures (e.g., Implicit Association Test), reflecting implicit evaluations, support this claim. Rydell et al. (2006) found a striking dissociation when they brief flashed either positive or negative words prior to presenting a photograph of a person was with behavioral information of the opposite valence was presented: IAT scores reflected the valence of the flashed words whereas rating scores reflected the opposite valence of the behavioral information. A recent study, however, suggests that this finding may not be replicable. Given its theoretical importance, we report two new replication attempts ($n = `r otm1_n`$ recruited in Belgium, Germany and the USA; $n = `r otm2_n`$ recruited in Hong Kong and the USA).

<@~{#intro-paragraph}

Are our explicit and implicit evaluations of an object or person always consistent with one another?
Or is it possible that we like a person explicitly but simultaneously dislike them implicitly?
One way to investigate this question is to compare two families of evaluative measures: direct measures (e.g., Likert scales) that assumedly elicit relatively more explicit, conscious, effortful, and controllable evaluations (hereafter explicit evaluations), on the one hand, and indirect measures [such as the Implicit Association Test [IAT]; @greenwald_measuring_1998] that assumedly elicit relatively more implicit, unconscious, effortless, and uncontrollable evaluations (hereafter implicit evaluations), on the other hand.
Indeed, several studies have shown dissociations between direct and indirect measures [see  @harmon-jones_what_2019]. 
Such evidence has been critical in supporting dual-process theories positing that explicit and implicit evaluations reflect different sets of attitudes that are acquired via two distinct processes.[^attitude_definition]

[^attitude_definition]:
By *attitude* we mean latent knowledge representations that underlie the behavioral expression of *evaluations* on direct and indirect measures [@cunningham_attitudes_2007].

~@>

An influential dual-process theory is the Systems of Evaluation Model [SEM; @mcconnell_systems_2014; @mcconnell_forming_2008; @rydellUnderstandingImplicitExplicit2006]. 
This theory assumes that implicit evaluations emerge from mental associations that develop without conscious awareness or control, from the co-occurrence of stimuli with valenced events.
For example, positive associations may develop simply because a person repeatedly wears a shirt in one’s favorite color.
In contrast, explicit evaluations are thought to reflect propositional representations that emerge from conscious, attention-demanding reasoning processes.
For example, negative propositions may develop as a result of learning that the person holds political opinions that clash with one’s own views.
Hence, under this theory, a double dissociation between direct and indirect measures of evaluation is expected, with the former reflecting only consciously formed propositions and the latter reflecting only unconsciously formed associations.

As a test of this model, @rydell_two_2006 contrasted two different learning pathways experimentally.
In the experiment, participants learned about an unfamiliar person called Bob.
Each trial started with a brief (25 ms) flash of a positive or negative word, not intended to be consciously registered by participants.
Then a photograph of Bob was presented alone for 250 ms before a positive or negative behavioral statement was added to the display.
The statement was clearly visible until participants made a guess as to whether the behavior was characteristic or uncharacteristic of Bob.
Participants immediately received feedback, which implied that Bob was a good or bad person.
Crucially, this behavioral information was always opposite in valence to the briefly flashed word.
In line with the predictions of the SEM, explicit evaluations of Bob, measured via self-report, reflected predominantly the valence of the behavioral information.
More intriguingly, implicit evaluations, measured via the IAT, reflected predominantly the valence of the words that had been briefly flashed prior to the photograph of Bob.

This finding has been influential in support of the SEM and other dual-process theories [e.g., @gawronskiAssociativePropositionalEvaluation2011]. 
However, beyond this prominent result, empirical evidence for dual evaluative learning processes remains weak overall [@corneilleAssociativeAttitudeLearning2018a]. 
The absence of compelling evidence that implicit evaluations emerge from unconsciously formed associations has allowed for a different, more parsimonious, account to be popularized: that both implicit and explicit evaluations reflect propositional knowledge [e.g., @de_houwer_propositional_2018]. 
Crucially, many prominent single-process propositional theories assume that propositional learning requires conscious awareness [@mitchellPropositionalNatureHuman2009]. 
As such, the result reported by @rydell_two_2006, where implicit evaluations reflected predominantly unconsciously formed associations, is particularly difficult to reconcile with these accounts.
Under most propositional theories, both direct (self-report) measures and indirect measures (such as the IAT) should reflect propositional knowledge that emerges from conscious, attention-demanding reasoning processes.

Given the theoretical issues at stake, a replication of the double dissociation reported by  @rydell_two_2006 is critical. 
If the double dissociation is replicated, such a result would lend credence to strong forms of dual-process theories positing that implicit and explicit evaluations reflect different types of (associative and propositional) representations that are acquired via different learning pathways.
Moreover, such a finding would provide evidence in favor of subliminal associative learning, a phenomenon for which current evidence is weak at best [@corneilleAssociativeAttitudeLearning2018a]. 
On the other hand, if the finding by @rydell_two_2006 does not replicate, and both direct and indirect measures are found to reflect the valence of the consciously processed behavioral information, such a result would strengthen confidence in single-process propositional theories of evaluation.
After all, these theories argue that both implicit and explicit evaluations largely reflect the same consciously formed propositions.

In two recent experiments, the double dissociation reported by @rydell_two_2006 did not replicate [@heycke_two_2018]. 
Instead, both direct and indirect measures consistently reflected the valence of the behavioral information.
At present, it is unclear whether these results point towards boundary conditions or call into question the replicability of the original study more generally.
This ambiguity is due to the fact that materials were translated into German and stimuli were presented for a duration different from the original study.
Here, we rigorously test the replicability of the double dissociation by closely adhering to the original procedure.
To ensure its informativeness, the current replication attempt was conducted jointly by an international collective of experts on evaluative learning and implicit measures.
Among the collaborators were the first author of the original study and authors of the previous replication attempts.
To explore the generality of our results, we collected data in multiple countries and languages.
A first, already concluded, experiment was conducted in Belgium, Germany and the USA.
In a second experiment, for which the data is yet to be collected, we will use the insights from the first experiment to adjust the procedure to closely replicate the psychological conditions of the original study.

@@

  # Recode, calculate indices, and filter data
  otm1_eval <- otm1_eval %>%
    mutate_at(vars(starts_with("Eval")), scale) %>%
    select(starts_with("Eval")) %>%
    rowwise %>%
    do(data.frame(Eval = mean(unlist(.)))) %>% 
    bind_cols(otm1_eval, .) %>%
    mutate(
      ParticipantNumber = factor(ParticipantNumber)
      , ValenceBlock = ifelse(ValenceBlock == 1, "Positive-negative", "Negative-positive") %>% factor
      , MeasureOrder = ifelse(MeasureOrder == 1, "Implicit-explicit", "Explicit-implicit") %>% factor
      , Block = factor(Block)
    )

@@

Because the procedural modifications made by @heycke_two_2018 may have caused the diverging results, we conducted a replication study using the unmodified experimental procedure of the original study.

@@

The first author of the original study verified that our materials and procedure faithfully reproduced the original.
The experiment was preregistered (https://osf.io/xe8au/) and data were collected at the University of Cologne (Germany), Ghent University (Belgium), and Harvard University (USA).
All data files, materials, and analysis scripts are available at https://osf.io/8m3xb/.
To give a vivid impression of the experimental procedure, an examplary video recording is available at https://osf.io/hmcfg/.

@@

The experimental procedure consisted of three components: a learning task, evaluation task, and recognition task.

As in the original study, the learning task was a modified version of the evaluative learning paradigm by @kerpelmanPartialReinforcementEffects1971. 
We briefly flashed a valent word followed by a longer presentation of a photograph of Bob together with a behavioral statement.
Presentation durations differed across labs due to the availability of different refresh rates of the CRT monitors (85 Hz at Harvard and 75 Hz at Ghent and Cologne).
In the following we will describe the setup of a trial with the presentation durations at a 75 Hz-refresh rate; deviating durations for a 85 Hz-refresh rate are given in brackets.

On each trial, a central fixation cross was displayed for 200 ms followed by a valent word flashed for 27 ms (24 ms; 2 frames).
The screen background was black and text was white and set in Times New Roman font. 
The briefly flashed word was immediately replaced by the photograph of Bob, which served as a backward mask.
Next, we provided behavioral information about Bob consisting of a behavioral statement and the additional information whether this behavior was characteristic or uncharacteristic of Bob.
The photograph of Bob was presented in the center of the screen for 253 ms (247 ms) before a behavioral statement was added underneath.
Participants’ task was to press the "c" (= "characteristic") or "u" (= "uncharacteristic") key to guess whether the behavioral statement was characteristic or uncharacteristic of Bob. 
After every guess, the photograph of Bob, the behavioral statement, and the key labels were replaced with either the word "Correct" displayed in green letters or the word "False" in red letters, displayed for 5000 ms.
Each trial ended with a blank screen presented for 1000 ms.

As the valence of briefly flashed words was manipulated within participants, they completed two 100-trial-blocks of the learning task.
Each block consisted of trials with either only positive or negative words and the order of the blocks was randomized.
The valence of the behavioral information was always opposite to the valence of the briefly flashed word.
In blocks with positive words, positive behavioral statements were uncharacteristic of Bob and negative statements were characteristic.
These contingencies were reversed in the blocks with negative words. 
We used 10 positive and 10 negative words; each of which was presented 10 times. 
For behavioral statements, we used 100 positive and 100 negative statements; 50 positive and 50 negative statements were randomly selected for the first block, the remaining statements were assigned to the second block. 
The order of briefly flashed words and behavioral information was randomized for each participant anew, whereas the order of blocks was counterbalanced across participants. 
A different photograph of Bob was randomly selected from six photographs of white males for each participant. 
The remaining five images were used in the implicit association test (see below). 
All materials were taken from the original study[^discrpaper], with the sole exception that briefly flashed words, behavioral statements, and instructions were translated to German and Dutch for use in Germany and Belgium.

[^discrpaper]:
The original manuscript lists the words  "love", "party", "hate", and "death" as examples for briefly flashed words. The words "hate" and "love", however, were neither used as briefly flashed words in the original, nor our replication studies.

After each block, we measured evaluations of Bob directly and indirectly using Likert-scale ratings and the IAT, respectively.
As in the original study, the order of the measures was the same for both blocks but counterbalanced across participants.

As direct measure of evaluation, we used three rating scales:
First, participants rated Bob's likableness on a 9-point slider with the anchors labelled *Very Unlikable* and *Very Likable*.
Next, again using 9-point sliders, they judged Bob on the dimensions *Bad--Good*, *Mean--Pleasant*, *Disagreeable--Agreeable*, *Uncaring--Caring*, and *Cruel--Kind*. 
Finally, they judged Bob on a 'feeling thermometer' by entering a number between 0 (*Extremely unfavorable*) and 100 (*Extremely favorable*).
Deviating from the original protocol, we collected rating scale responses as part of the computer task rather than using a paper-pencil questionnaire.

As indirect measure of evaluation, we used an IAT. 
Participants initially completed two types of training blocks with 20 trials each to familiarize themselves with the task.
In one block, images of Bob and other white men had to be classified as Bob vs. not-Bob; in another block, positive and negative words had to be classified as positive vs. negative. 
In a subsequent critical block with 40 trials we intermixed the two classification tasks: 
Participants used one key to respond to both the images of Bob and negative words; they used another key to respond to images of other white men and positive words.
After the first critical block, participants completed another training block with 20 trials of Bob vs. not-Bob with reversed key position and afterwards a second critical block with 40 trials with the reversed key mapping compared to the first critical block. 
It was counterbalanced whether participants completed the IAT as described above or with key mappings in reversed order [for a detailed description see @heycke_two_2018, p. 1712].
We instructed participants to respond quickly without making too many errors.
In case of erroneous responses we displayed a red X as feedback and instructed participants to quickly correct their response to start the next trial.

Following the first round of evaluations, participants completed the second learning block and again evaluated Bob directly and indirectly.
After the second round of evaluations, participants completed a surprise recognition test for the briefly flashed words.
We presented 40 words in random order on a computer screen.
Half of the words were the briefly flashed words from the learning task, the other half were new distractor words.
We informed participants that 20 words were flashed briefly during the learning task, asked them to select the briefly flashed words from the list, and encouraged them to guess if they did not know the correct answer.
Participants could only proceed with the experiment once they had selected exactly 20 words.

The experiment ended with a demographic questionnaire (age, field of study/profession, gender, goal of the experiment, and comments).
Our procedure was identical to the original procedure, with the exception that participants completed self-reported evaluations and the recognition task at the computer rather than using paper and pencil.
In Belgium and Germany, we furthermore used Dutch and German translations of the original material.
The procedure took approximately 50 minutes to complete.

### Data analysis {#data-analysis1}

<@~{#score-calculation}

In keeping with the original analysis strategy, we calculated composite rating scores and IAT scores as direct and indirect measures of evaluation.
Rating scores were the average of the three $z$-standardized Likert-scale responses.
To calculate IAT scores we logarithmized all response times after winsorizing responses faster than 300 ms or slower than 3,000 ms.
IAT scores were the difference of mean transformed response times for blocks which combined Bob and negative words and blocks which combined Bob and positive words.
Thus, for rating and IAT scores larger values indicate a more positive evaluation of Bob.

~@>

How to statistically assess the success of a replication attempt is subject of current debate [e.g., @fabrigar_conceptualizing_2016; @simonsohn_small_2013; @verhagen_bayesian_2014]. 
Whether a pattern of results has been replicated is challenging to measure directly if the to-be-replicated pattern consists of more than two cells of a factorial design.
One elegant approach is to instantiate a pattern of mean differences (i.e., the rank order of means), predicted by a theory or observed in a previous study, as order constraints in a statistical model [e.g., @hoijtink_informative_2012; @rouder_theories_2018].
With the model in hand, replication success can be quantified as predictive accuracy of this model relative to a competing model, such as a null model or an encompassing unconstrained model [e.g., @rouder_theories_2018].

Based on previously reported results, there are two competing predictions for the current paradigm:
(1) @rydell_two_2006 reported that across both learning blocks ratings scores were congruent with the behavioral information about Bob, whereas IAT scores were incongruent with the behavioral information ($\mathcal{H}_\textrm{Two minds}$).
(2) In contrast, @heycke_two_2018 observed a consistent pattern for rating scores and IAT scores; both measures were congruent with the behavioral information ($\mathcal{H}_\textrm{One mind}$).
We considered two additional predictions: no effect of the manipulation ($\mathcal{H}_\textrm{No effect}$) and the all-encompassing prediction of any outcome ($\mathcal{H}_\textrm{Any effect}$).
If, of all predictions considered, our results are best described by the prediction of no effect, our experimental manipulations failed.
The prediction of any effect reflects the possibility that we may observe an entirely unexpected outcome that is neither in line with the results reported by @rydell_two_2006 or @heycke_two_2018.

We implemented all predictions as order (or null) constraints in an ANOVA model with default (multivariate) Cauchy priors [$r = 0.5$ for fixed effects and $r = 1$ for random participant effects, see SOM for details; @rouder_default_2012; @rouder_theories_2018].
To simplify the presentation of the Bayesian model comparison results, we collapsed data across valence orders such that we always contrasted blocks where the behavioral information was positive with those where it was negative.
Thus, for both rating and IAT scores positive difference indicate that evaluations are congruent with the valence of the behavioral information, whereas negative values indicate that evaluations are congruent with the valence of the briefly flashed words.
We assessed the relative predictive accuracy of these models by Bayesian model comparisons using Bayes factors.
Note that comparisons of models where one model is a special order-constrained case of the other are asymmetric.
Consider the example of $\mathcal{H}_\textrm{One mind}$, which is a special case of $\mathcal{H}_\textrm{Any effect}$.
If the data are perfectly consistent with $\mathcal{M}_\textrm{One mind}$, they are inevitably also perfectly consistent with $\mathcal{M}_\textrm{Any effect}$.
In this case $\mathcal{M}_\textrm{One mind}$ will be favored by the Bayes factor because $\mathcal{M}_\textrm{One mind}$ makes a more specific prediction---it predicts that 3/4 of the outcomes predicted by $\mathcal{M}_\textrm{Any effect}$ are impossible, Figure \ \@ref(fig:otm1-bayesian-replication-analysis-plot)A.
The degree to which the order-constrained model is more specific (more parsimonious) places an upper bound on the Bayes factor in its favor.
On the other hand, there is no such bound on the Bayes factor in favor of the unconstrained model if the data are inconsistent with the order constraint---that is, the data fall outside of the predictive space deemd possible by the order-constrained model.
It follows that $\mathrm{BF}_{\mathcal{M}_\textrm{One mind}/\mathcal{M}_\textrm{Any effect}} \in [0, 4]$ because $\mathcal{M}_\textrm{One mind}$ limits its predictions to 1/4 of those of $\mathcal{M}_\textrm{Any effect}$.
To guide their interpretation, we report the theoretical bounds on the reported Bayes factors alongside our results where applicable.
Finally, we tested whether recognition memory accuracy using a one-tailed Bayesian $t$ test with default Cauchy prior [$r = \sqrt2/2$; @rouder_bayesian_2009].

To facilitate comparisons with previously reported statistics, we also conducted the frequentist analyses described by @rydell_two_2006.
To ensure that our conclusions about indirectly measured evaluations are robust to stimulus effects, we supplemented the ANOVA analysis of IAT scores by a frequentist linear mixed model analysis, see SOM.
We used `r cite_r("r-references.bib", pkgs = c("papaja", "afex", "emmeans", "BayesFactor"), withhold = FALSE)` for all our analyses.

@@

We set out to collect $n = 50$ participants at each location (N = 150).
We recruited `r otm1_n + otm1_n_excluded` participants (aged `r paste(range(otm1_mem$Age), collapse = "-")` years, $M = `r mean(otm1_mem$Age)`$; `r sum(grepl("female", otm1_mem$Sex)) / nrow(otm1_mem) * 100`% female, `r printnum(sum(grepl("nonbinary", otm1_mem$Sex)) / nrow(otm1_mem) * 100)`% nonbinary; see supplementary online material [SOM] for details); `r printnum(otm1_n_excluded, numerals = otm1_n_excluded > 10)` participants were excluded due to technical failures<!-- or noncompliance with instructions-->.
Hence, the reported results are based on data from `r otm1_n` participants.
We compensated all participants with either € 8/10 (Cologne/Ghent), or partial course credit (Cologne/Harvard).

### Statistical power

```{r rydell-effects, cache = TRUE}
rydell_pes <- 0.1
rydell_f <- 4.78
rydell_n <- 50
rydell_dz <- pes_to_dz(rydell_pes, rydell_n, n_groups = 2)

heycke_t <- c(2.54, 1.45)
heycke_n <- c(28, 26)
heycke_dz <- heycke_t / sqrt(heycke_n)
heycke_pes <- dz_to_pes(heycke_dz, heycke_n, n_groups = 1)
```

```{r otm1-power, cache = TRUE}
otm1_power <- function(f, rm_cor, n = 76, alpha = 0.05) {
  contrast_code <- tibble::tibble(
    a = rep(contr.sum(2), 2)
    , b = rep(contr.sum(2), each = 2)
    , x = a*b * -1
  )

  mu <- with(
    contrast_code
    , f * x +
      - 0.25 * a
  )
